Commutative Character Sheaves

### 2015-11-16 followup.tex
Lots of work on Section 4, but all very rough. Out of time for now...

Next, we broaden our scope further to encompass smooth group schemes $G$ that are not necessarily commutative.
In this paper, we do not consider Lusztig's geometrization of arbitrary characters of the non-abelian group $G(k)$, when $G$ is reductive,
using character sheaves \cite{lusztig:85a}*{Def.~2.10}, but focus on the case of $1$-dimensional characters
$\Hom(G(k), \EEx)$ assuming only that $G$ is smooth. The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more linear character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We define a category $\CCS(G)$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to \Hom(G\ab(k), \EEx)$
is an isomorphism.  In fact, pullback along the quotient $q : G \to G\ab$ defines an equivalence of categories
$\CCS(G\ab) \to \CCS(G)$.  Note that the functor $\CCS(G) \to \CS(G)$ is not essentially surjective, missing
the kinds of linear character sheaves highlighted by Kamgarpour.  Moreover, $\Hom(G\ab(k), \EEx)$ surjects
onto $\Hom(G(k), \EEx)$, with a kernel described by Galois cohomology.

@@

\todo{We should give the relationship between $\CCSiso{G}$ and $G(k)^*$ here as a corollary.  Also, include the automorphism group of a commutative character sheaf.}
\todo{Is this what you had in mind?}

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the generic data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from the generic data. 
Let us refer to the types arising in this manner as {\it generic types}.
In this section we show that every generic type is geometrizable, in a sense which will be made clear below.

@@

As explained in \cite{yu:models}, a generic type may be apprehended as a pair $(\underline{G}(\OK), \rho)$, where $\underline{G}$ is an affine smooth group scheme of finite type over $\OK$ and where $\rho$ is a representation of $\underline{G}(\OK)$.
The group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\prod_{i=0}^d \underline{J}^i(\OK) = \underline{G}(\OK),\] where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$, and such that $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, for $j=1, \ldots, d$;
\item affine smooth group schemes $\underline{G}^{j}$, for $j=1, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = \prod_{i=0}^{j} \underline{J}^i(\OK),
\] 
where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$ and $\underline{G}^{j}(\OK)$ with its image under $\underline{G}^j \to \underline{G}$.
\end{enumerate}
The representation $\rho$ of $\underline{G}(\OK)$ is build from the following data:
\begin{enumerate}
\item
	a representation $\sigma_0$ of $\underline{G}_s\red(\Fq)$;
\item
	a representation $\epsilon_j$ of $(\underline{G}^{j-1})_s\red(\Fq)$, for each $j=1,\ldots, d$;
\item
	a representation $\eta_j$ of $H^j(\Fq)$, where $H^j$ is Heisenberg group arising as a quotient $(\underline{J}^{j})_s \to H^i$, for each $j=1,\ldots, d$;
\item
	a linear quasicharacter $\phi_{j-1}$ of $\underline{G}^{j-1}(\OK)$, for $j=1,\ldots, d$;
\item
	a linear quasicharacter $\tau_d$ of $\underline{G}(\OK)$.
\end{enumerate}
From this data, the representation $\rho$ is built as follows.
\begin{enumerate}
\item 
	Let $\tau_0$ be the representation of $\underline{G}(\OK)$ obtained by pulling back $\sigma_0$ along 
\[\underline{G}(\OK) \to \underline{G}_s(\Fq) \to \underline{G}_s\red(\Fq).\]
\item 
	Let $\dot{\epsilon}_j$ be the representation of $\underline{G}^{j-1}(\OK)$ obtained by pulling back $\epsilon_j$ along 
\[\underline{G}^{j-1}(\OK) \to \underline{G}^{j-1}_s(\Fq) \to (\underline{G}^{j-1})_s\red(\Fq),\]
for $j=1,\ldots, d-1$.
\item
	Let $\dot{\eta}_j$ be the representation of $\underline{J}^j(\OK)$ obtained by pulling back $\eta_j$ along 
\[\underline{J}^{j}(\OK) \to \underline{J}^{j}_s(\Fq) \to H^j(\Fq),\]
for $j=1,\ldots, d-1$.
\item
	Using the decomposition $\underline{G}^j(\OK) = \underline{G}^{j-1}(\OK) \times \underline{J}^j(\OK)$, consider the representation $(\phi_{j-1}\otimes \dot{\epsilon}_j)\times \dot{\eta}_j$ of $\underline{G}^{j}(\OK)$, for $j=1,\ldots, d-1$. 
	Since $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, this determines a representation $\tau_j$ of $\underline{G}(\OK)$, for $j=1,\ldots, d-1$. 
\item
	Then $\rho = \tau_0 \otimes \cdots \tau_d$.
\end{enumerate}

Regarding this construction of the type $(\underline{G}(\OK),\rho)$, Yu points out that,
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:models}
\end{quotation}
%

@@

The objective of this section is to push Yu's observation further to prove the following result.

\begin{proposition}\label{prop:types}
Let $(\underline{G}(\OK),\rho)$ be a generic type.
Set $G = \Gr(\underline{G})$, the Greenberg transform of $\underline{G}$; recall that this is a group scheme over $\Fq$.
There is a virtual admissible $\ell$-adic sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}

@@

Now, let $(A,\varphi)$ be a Frobenius-stable character sheaf on $\underline{G}_s\red$.
Let $(A,\varphi)$ be the Frobenius-stable constructible sheaf on $\underline{G}_s$ obtained by pull-back.
For each $j=1,\ldots, d-1$, 
let $(A^j, \varphi^j)$ be a character sheaf on $(\underline{G}^{j-1}_s)\red$, in the sense of [Lusztig, CSDG].
Let $(C_1^j, \phi^j)$ be the Frobenius-stable constructible sheaf on $\underline{G}^{j-1}_s$ obtained by pull-back..

For each $j=1,\ldots, d-1$,
let $H^j$ be a unipotent group appearing as a quotient $\underline{J}^j_s \to H^j$.
Let $B^j$ be a character sheaf on $H^j$, in the sense of [Boyarchenko].
Let $(D_1^j,\varphi^j)$ be the Frobenius-stable constructible sheaf on $\underline{J}^{j}_s$ obtained by pull-back.

Let $n$ be a positive integer and
for each $j=1,\ldots, d-1$, let $\cs{L}_n^{j-1}$ be a commutative quasicharacter sheaf on $G_n^{j-1}\ceq \Gr_n(\underline{G}^{j-1})$.

Observe that $\underline{G}^{j-1}_s = \Gr_1(\underline{G}^{j-1})$ 
and $\underline{J}^{j}_s = \Gr_1(\underline{G}^{j-1})$. 
Let $(\cs{C}_n^j,\phi^j)$ be the admissible sheaf on $G_n^{j-1} \ceq \Gr_n(\underline{G}^{j-1})$ obtained by pulling-back $C_1$ along $G_n^{j-1} \to \underline{G}^{j-1}_s$.
Let $(\cs{D}_n^j,\phi^j)$ be the admissible sheaf on $J_n^{j-1} \ceq \Gr_n(\underline{J}^{j-1})$ obtained by pull-back along $J_n^{j-1} \to \underline{J}^{j-1}_s$.

Set $\underline{N}^j = \prod_{i=j}^d \underline{J}^i$ and let $\underline{G} \to \underline{M}^j$ be the morphism . . . .
Then $\underline{M}^j = \underline{G}^{j-1} \underline{J}^j$ . . . 
%
Now, consider the following sheaf on $M_n^j \ceq \Gr_n(\underline{M}^j)$:
\[
\cs{E}_n^j \ceq (\cs{L}_n^{j-1}\otimes \cs{C}_n^j) \boxtimes \cs{D}_n^{j}
\]
Let $\cs{F}_n^j$ be the pullback along $G_n :=\Gr_n(\underline{G}) \to \Gr_n(\underline{M}^j):= M_n^j$.
Finally, set $\cs{F}_n = \otimes_i \cs{F}^i_n$.

>>>>

Next, we broaden our scope further to encompass smooth group schemes $G$ that are not necessarily commutative.
In this paper, we do not consider Lusztig's geometrization of arbitrary characters of the non-abelian group $G(k)$, where $G$ is reductive,
using character sheaves \cite{lusztig:85a}*{Def.~2.10}, but focus on the case of $1$-dimensional characters
$\Hom(G(k), \EEx)$ assuming only that $G$ is smooth. The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more linear character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We define a category $\CCS(G)$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to \Hom(G\ab(k), \EEx)$
is an isomorphism.  In fact, pullback along the quotient $q : G \to G\ab$ defines an equivalence of categories
$\CCS(G\ab) \to \CCS(G)$.  Note that the functor $\CCS(G) \to \CS(G)$ is not essentially surjective, missing
the kinds of linear character sheaves highlighted by Kamgarpour.  
Since $\Hom(G\ab(k), \EEx)$ surjects
onto $\Hom(G(k), \EEx)$, with a kernel described by Galois cohomology, this paper shows that, for every linear character $\chi : G(k) \to \EEx$ there is a commutative character sheaf $\cs{L}$ on $G$ such that $\TrFrob{G}(\cs{L})\circ q = \chi$.

@@

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from the data. 
Let us refer to the types arising in this manner as {\it Yu types}.
In this section we show that every Yu type is geometrizable, in a sense which will be made clear below.

@@

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
One of the main results of \cite{yu:models} is that there is a canonical affine smooth group scheme $\underline{G}$ of finite type over $\OK$ such that $\,^\circ K = \underline{G}(\OK)$ \cite{yu:models}*{Prop 10.2}.
As explained in \cite{yu:models}*{\S 10.4}, the
the group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\prod_{i=0}^d \underline{J}^i(\OK) = \underline{G}(\OK),\] where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$, and such that $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, for $j=1, \ldots, d$;
\item affine smooth group schemes $\underline{G}^{j}$, for $j=1, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = \prod_{i=0}^{j} \underline{J}^i(\OK),
\] 
where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$ and $\underline{G}^{j}(\OK)$ with its image under $\underline{G}^j \to \underline{G}$.
\end{enumerate}

The representation $\,^\circ \rho$ of $\underline{G}(\OK)$ is build from the following data, as explained in \cite{yu:models}*{\S 10.5}:
\begin{enumerate}
\item
	a representation $\sigma$ of $\underline{G}_s\red(\Fq)$;
\item
for $j=1,\ldots, d-1$, 
\begin{enumerate}
\item
a representation $\eta_j$ of $H^j(\Fq)$, where $H^j$ is Heisenberg group arising as a quotient $(\underline{J}^{j})_s \to H^i$,
\item
a representation $\epsilon_j$ of $(\underline{G}^{j-1})_s\red(\Fq)$
\item
a linear quasicharacter $\phi_{j-1}$ of $\underline{G}^{j-1}(\OK)$, for $j=1,\ldots, d$;
\end{enumerate}
\item
	a linear quasicharacter $\tau_d$ of $\underline{G}(\OK)$.
\end{enumerate}
From this data, the representation $\,^\circ \rho$ is built as follows.
\begin{enumerate}
\item 
	Let $\tau_0$ be the representation of $\underline{G}(\OK)$ obtained by pulling back $\sigma$ along 
\[\underline{G}(\OK) \to \underline{G}_s(\Fq) \to \underline{G}_s\red(\Fq).\]
\item For $j=1,\ldots, d-1$, 
\begin{enumerate}
\item
	Let $\dot{\eta}_j$ be the representation of $\underline{J}^j(\OK)$ obtained by pulling back $\eta_j$ along 
\[\underline{J}^{j}(\OK) \to \underline{J}^{j}_s(\Fq) \to H^j(\Fq),\]
for $j=1,\ldots, d-1$.
\item 
	Let $\dot{\epsilon}_j$ be the representation of $\underline{G}^{j-1}(\OK)$ obtained by pulling back $\epsilon_j$ along 
\[\underline{G}^{j-1}(\OK) \to \underline{G}^{j-1}_s(\Fq) \to (\underline{G}^{j-1})_s\red(\Fq),\]
for $j=1,\ldots, d-1$.
\item
	Using the decomposition $\underline{G}^j(\OK) = \underline{G}^{j-1}(\OK) \times \underline{J}^j(\OK)$, consider the representation 
\[
	(\phi_{j-1}\otimes \dot{\epsilon}_j)\times \dot{\eta}_j
\]
 of $\underline{G}^{j}(\OK)$, for $j=1,\ldots, d-1$. 
	Since $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, this determines a representation $\tau_j$ of $\underline{G}(\OK)$, for $j=1,\ldots, d-1$. 
\end{enumerate}
\item
	Then $\,^\circ \rho = \tau_0 \otimes \cdots \tau_d$.
\end{enumerate}

Regarding this construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:models}
\end{quotation}
%
The objective of this section is to push this observation further by geometrizing the representation $\,^\circ \rho$ of $\underline{G}$.

@@

Let $(\underline{G}(\OK),\,^\circ \rho)$ be a Yu type.
Set $G = \Gr(\underline{G})$, the Greenberg transform of $\underline{G}$.
Recall that $G$ is a group scheme over $\Fq$ and comes equipped with

\begin{proposition}\label{prop:types}
Let $(\underline{G}(\OK),\rho)$ be a Yu type.
With notation as above, suppose $\pi_0(\underline{G}_s\red)$ is cyclic and that $\pi_0((\underline{G}^{j-1})_s\red)$ is cyclic for $j=1,\ldots, d-1$.
Then there is a virtual admissible $\ell$-adic sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}

@@

Introduce morphisms $\underline{G}\to \underline{M}^j$ and $\underline{M}^{j} \to \underline{M}^{j-1}$ and $\underline{M}^j \to \underline{J}^j$.
\[
\begin{tikzcd}
\underline{J}^j \arrow{r} & \underline{G}^{j} \arrow{r} 
& \underline{G} \arrow{r}  
& \underline{M}^{j} \arrow{d} \arrow{r} & \underline{J}^j \\
& \arrow{u} \underline{G}^{j-1} 
&
& \underline{M}^{j-1}
&
\end{tikzcd}
\]
Interpret 
\begin{enumerate}
\item
	$\epsilon_j$ as a representation of $(\underline{M}^{j-1})_s\red(\Fq)$, for each $j=1,\ldots, d$;
\item
	$\phi_{j-1}$ as a linear quasicharacter of $\underline{M}^{j-1}(\OK)$, for $j=1,\ldots, d$;
\end{enumerate}
so that
\begin{enumerate}
\item 
	$\dot{\epsilon}_j$ is the representation of $\underline{M}^{j-1}(\OK)$ obtained by pulling back $\epsilon_j$ along 
\[\underline{M}^{j-1}(\OK) \to \underline{M}^{j-1}_s(\Fq) \to (\underline{M}^{j-1})_s\red(\Fq),\]
for $j=1,\ldots, d-1$.
\item
	$\dot{\eta}_j$ is a representation of $\underline{J}^{j}(\OK)$ obtained by pulling back  . . . .  $j=1,\ldots, d-1$;
\item
	$\tau_j$ is the representation of $\underline{G}(\OK)$ obtained by pulling back  $\phi_{j-1}$ and $\dot{\epsilon}_j$ and $\dot{\eta}_j$ from $\underline{M}^{j}(\OK)$ along $\underline{G}(\OK)\to \underline{M}^j(\OK)$, for $j=1,\ldots, d-1$. 
\end{enumerate}

\[
\begin{tikzcd}
\underline{G} \arrow{r}  
& \underline{M}^{j} \arrow{d} \arrow{r} & \underline{J}^j \\
\underline{G}_s \arrow{u} \arrow{d}
& \underline{M}^{j-1}
 & (\underline{J}^j)_s \arrow{u} \arrow{d} \\
\underline{G}_s\red & (\underline{M}^{j-1})_s \arrow{u} \arrow{d} & H^j \\
& (\underline{M}^{j-1})_s\red & 
\end{tikzcd}
\]

Now, we introduce the tools needed to geometrize all the ingredient representations.
\begin{enumerate}
\item
Let $(A_0,\varphi)$ be a Frobenius-stable character sheaf on $\underline{G}_s\red$, in the sense of [Lusztig].
Let $(A,\varphi)$ be the Frobenius-stable constructible sheaf on $\underline{G}_s$ obtained by pull-back.
\item For $j=1,\ldots, d-1$, 
\begin{enumerate}
\item
Let $(B_0^{j},\varphi)$ be a Frobenius-stable character sheaf on $H^j$ in the sense of [Boyarchenko].
Let $(B^{j},\varphi)$ be the Frobenius-stable constructible sheaf on $\underline{J}^j_s$ obtained by pull-back.
\item
Let $(C_0^{j-1},\varphi)$ be a Frobenius-stable character sheaf on $(\underline{M}^{j-1})_s\red$ in the sense of [Lusztig].
Let $(C^{j-1},\varphi)$ be the Frobenius-stable constructible sheaf on $\underline{M}^{j-1}_s$ obtained by pull-back.
\item
Let $\cs{E}_n$ be a commutative quasicharacter sheaf on $M_n^{j-1}$.
\end{enumerate}
\item
Let $\cs{L}_n$ be a commutative quasicharacter sheaf on $G_n$.
\end{enumerate}
\[
\begin{tikzcd}
G \arrow{r}  \arrow{d}
& M^{j} \arrow{r} \arrow{d}
& J^j \arrow{d} \\
G_n \arrow{r}  \arrow{d}
& M_n^{j} \arrow{r} \arrow{d}
& J_n^j \arrow{d} \\
\underline{G}_s \arrow{d}
& \underline{M}^{j-1} \arrow{d}
& (\underline{J}^j)_s  \arrow{d} \\
\underline{G}_s\red 
& (\underline{M}^{j-1})_s \arrow{d} & H^j \\
& (\underline{M}^{j-1})_s\red 
& 
\end{tikzcd}
\]

We may now encorporate all these ingredients into one sheaf on $G_n$.
\begin{enumerate}
\item
Let $\cs{A}_n$ be sheaf on $G_n$ produced from $A$ by pullback along $G_n \to \underline{G}_s$.
\item
For $j=1, \ldots, d-1$,
\begin{enumerate}
\item let $\cs{B}^j$ be the sheaf on $M^j_n$ ...
\item let $\cs{C}^j$ be the sheaf on $M^j_n$ ...
\item let $\cs{E}^j$ be the sheaf on $M^j_n$ ...
\end{enumerate}
\item Let $\cs{L}$ be the sheaf on $G_n$ as above...
\end{enumerate}
Combine: $B$, $C$ and $E$ into $D$
\[
\cs{A} \otimes \cs{D}^1 \otimes \cdots \otimes \cs{D}^{d-1} \otimes \cs{L}.
\]
